# Setting up for this course

## Phase 1: Creating Your Repository from Template

```{r, out.width = "100%", fig.alt = "Welcome to Reproducibility Capstone. This course is brought to you by the I T C R Training Network.", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1ExTZsKDHKM0fIi0_tB80qd8cHvqjEDyNJp-OsnvZ19o/edit?slide=id.g360ed975019_0_188#slide=id.g360ed975019_0_188")
```

### Step 1: Access the Course Template

<input type="checkbox"> Navigate to the [course sandbox template repository](https://github.com/fhdsl/capstone-sandbox)<br/>
<input type="checkbox"> Look for the <div class = "github_button"> <a href="https://github.com/new?template_name=capstone-sandbox&template_owner=fhdsl"> Use this template</a></div> green button at the top of the repository page<br/>

::: {.notice}
Important: Do NOT fork the repository - you must use the template option
:::

### Step 2: Configure Your New Repository

<input type="checkbox"> Click <div class = "github_button"> <a href="https://github.com/new?template_name=capstone-sandbox&template_owner=fhdsl"> Use this template</a></div> and select the "Create a new repository" option (if needed).<br/>
<input type="checkbox"> Fill out the repository creation form:<br/>
  • <input type="checkbox"> Include all branches: Leave **unchecked** (we only need the main branch)<br/>
  • <input type="checkbox"> Repository name: Choose a meaningful name (e.g., `your-username-capstone-sandbox`)<br/>
  • <input type="checkbox"> Description: Add a brief description of your capstone project<br/>
  • <input type="checkbox"> Visibility: Set to **Public** (required for the automated evaluation system to work)<br/>


### Step 3: Create Your Repository

<input type="checkbox"> Click <div class = "github_button"> Create repository </div><br/>
<input type="checkbox"> Wait for GitHub to finish creating your repository and verify completion <br/>
  • You'll be redirected to your new repository page<br/>
  • <input type="checkbox"> Verify: The repository should show your username as the owner and contain all the template files<br/>

## Phase 2: Local Setup and Cloning

For this course you may use whatever git client (GitHub Desktop, GitKraken, etc) you are most comfortable with, or no git client at all.
If you need any recap on how to file pull requests remember you can return to the other courses in this series for reference. Particularly the [Advanced Reproducibility course](https://jhudatascience.org/Adv_Reproducibility_in_Cancer_Informatics/using-version-control-with-github.html)

### Step 4: Clone Your Repository {.tabset}

#### Option A: Using Command Line

1. Copy the repository URL:<br/>
   • <input type="checkbox"> On your repository page, click the <div class = "github_button"> Code </div> green button <br/>
   • <input type="checkbox"> Copy the HTTPS URL (should look like: `https://github.com/yourusername/your-repo-name.git`) <br/>

2. Clone the repository, using the HTTPS URL you've just copied:
   ```bash
   # Navigate to where you want to store the project
   cd ~/Desktop  # or your preferred location

   # Clone the repository
   git clone https://github.com/yourusername/your-repo-name.git

   # Navigate into the repository
   cd your-repo-name
   ```

#### Option B: Using GitHub Desktop

1. Open GitHub Desktop
2. Click "Clone a repository from the Internet"
3. Select your repository from the list or paste the URL you copied
4. Choose a local path where you want to store the repository
5. Click "Clone"

## Phase 3: Branch Creation and Workflow

### Step 6: Create Your First Assignment Branch

You may name your working branch whatever you like for these assignments. The examples/instructions use the name `assignment-1-dockerfile`.

#### Command Line Method:

```bash
# Create and switch to a new branch for Assignment 1
git checkout -b assignment-1-dockerfile

# Verify you're on the new branch
git branch
# The asterisk (*) should be next to "assignment-1-dockerfile"
```

#### GitHub Desktop Method:
1. Click on the "Current branch" dropdown (shows "main")
2. Click "New branch"
3. Name your branch: `assignment-1-dockerfile`
4. Click "Create branch"

## Your Setup is Complete.... What are the Next Steps?

This setup process creates the foundation for all your capstone assignments. The first two assignments will follow a similar pattern:

create branch → make changes → commit → push → open PR → receive evaluation → iterate if needed.

```{r out.width="100%", echo = FALSE, fig.alt = "For each assignment, create a new branch. Then make edits to accomplish the task of the assignment. Optionally you may be able to test those edits locally (e.g., building a docker image locally). Git commit and push those edits and open a pull request (or check an existing pull request) to receive evaluation from automated assignment checks. Interpret the results. If you were not successful, troubleshoot and iterate this process as needed, checking your pre-existing pull request for new evaluation results after making additional changes. Or if you were successful in your assignment, submit the verification code to your coursera quiz (or move on to the next section if you are not using coursera or some certification method to take this course)"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1ExTZsKDHKM0fIi0_tB80qd8cHvqjEDyNJp-OsnvZ19o/edit?slide=id.g368818bb4f3_0_353#slide=id.g368818bb4f3_0_353")
```

The automated evaluation system depends on this workflow, so following these steps precisely is crucial for getting your validation codes!

Note that the validation codes are really only needed if you are participating in this capstone as a part of the Coursera or Leanpub platforms (for certification). If you are not taking this course for certification, there is no structure in place to verify the validation code you receive.


### Reminders of Best Practices for Success while Working with GitHub

- Use one branch per assignment - keeps work organized
- Write descriptive commit messages - helps track your progress
- Regularly push your work - protects your work and enables evaluation
- Use clear titles and descriptions for pull requests - will help you and your collaborators know what the work in the branch is
- Don't merge your pull request before you get the validation code - outside of this assignment, wait for approval from a team member or collaborator
- Monitor automated feedback - respond to or handle issues identified by evaluation promptly

## Troubleshooting Common Issues

### "Repository not found" when cloning:
- Check that the repository is public
- Verify the URL you've copy/pasted is correct
- Ensure you have access to the repository

### "Permission denied" when pushing:
- Check that you're pushing to your own repository (not the template)
- Verify your Git credentials are set up correctly
- Try using a personal access token instead of password

### Automated evaluation not running:
- Ensure your pull request is against the main branch
- Check that required files are in the correct locations
- Verify the repository is public
