# Making a GitHub Secret

```{r, out.width = "100%", fig.alt = "Learning Objectives: Recognize the purpose of GitHub secrets; Explain how to store your GitHub secrets in a secure way; Create a personal access token (PAT) and store it in your GitHub repository; Test your GitHub secret", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1ExTZsKDHKM0fIi0_tB80qd8cHvqjEDyNJp-OsnvZ19o/edit?slide=id.g360ed975019_0_0#slide=id.g360ed975019_0_0")
```

## Phase 0: Understanding GitHub Secrets

### What Are GitHub Secrets?

[GitHub Secrets](https://docs.github.com/en/actions/concepts/security/about-secrets) are encrypted environment variables stored securely in your repository. They allow you to store sensitive information (like usernames, API keys, tokens, passwords) without exposing them in your code. Secrets are only accessible to GitHub Actions during workflow runs and appear as `*` in logs.

This assignment teaches you a fundamental skill for secure automation in research environments

### Why Personal Access Tokens Matter

[Personal Access Tokens (PATs)](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/managing-your-personal-access-tokens) provide secure authentication for GitHub Actions. They're an alternative to passwords for GitHub authentication. They allow automated workflows to perform actions that require permissions (like creating issues, pushing to repositories, etc.). In research contexts, PATs enable automated data sharing, publication workflows, and collaborative tools

::: {.warning}
Important: PATs should be treated like passwords - never share them publicly
:::

Often, and in this assignment, you will begin by generating a personal access token in GitHub settings and [storing that as a repository secret](https://docs.github.com/en/actions/security-for-github-actions/security-guides/using-secrets-in-github-actions#creating-secrets-for-a-repository) that can be [utilized within workflows/GitHub Actions](https://docs.github.com/en/actions/security-for-github-actions/security-guides/using-secrets-in-github-actions#using-secrets-in-a-workflow). But, as mentioned above, GitHub secrets do not always have to be personal access tokens.

### Best Practices and Security Considerations

#### Token Management

- Use descriptive names for tokens to track their purpose
- Set appropriate expiration dates - shorter is more secure
- Regularly audit and rotate tokens in production environments
- Revoke unused tokens to minimize security exposure

#### Secret Storage Best Practices

- Use specific, descriptive names for secrets
- Apply principle of least privilege - only grant necessary permissions
- Never log or print secrets in your workflows
- Regularly review who has access to repository secrets

#### Research Context Applications

- API keys for data sources (e.g., cancer databases, genomic repositories)
- Authentication for computing clusters or cloud services
- Credentials for automated publication or preprint systems
- Tokens for collaborative tools and notification systems

## Phase 1: Creating Your Personal Access Token

### Step 1: Navigate to GitHub Token Settings

<input type="checkbox"> Go to your GitHub profile (click your profile picture in the top right)<br/>
<input type="checkbox"> Click "Settings" from the dropdown menu (you will see a gear icon `r config::get("settings")`)<br/>
<input type="checkbox"> Scroll down to "Developer settings" in the left sidebar (at the bottom)<br/>
<input type="checkbox"> Click "Personal access tokens" `r config::get("token")` <br/>
<input type="checkbox"> Select "Tokens (classic)" from the submenu<br/>

```
Navigation path: Profile → Settings → Developer settings → Personal access tokens → Tokens (classic)
```

### Step 2: Generate New Token

<input type="checkbox"> Click "Generate new token"<br/>
<input type="checkbox"> Select "Generate new token (classic)"<br/>
<input type="checkbox"> GitHub may ask for your password or a 2-factor authenticator (2FA) code - enter it to continue<br/>
<input type="checkbox"> You'll be taken to the token creation page<br/>

### Step 3: Configure Your Token

<input type="checkbox"> Add a descriptive note:<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • Use something like "Secret Storage Assignment"<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • This helps you remember what this token is for later<br/>

<input type="checkbox"> Set expiration: <br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • For this assignment, you can set it to 30 days or custom<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • In real research, consider the security vs. convenience trade-off<br/>

<input type="checkbox"> Select scopes (permissions): <br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • Check "repo" - this gives full repository access<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • For this assignment, "repo" scope is sufficient<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • If using a GitHub secret in a GitHub Action/workflow, you may need to select "workflow" as well.<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; • In production, use the minimum required permissions<br/>

```{r echo=FALSE, out.width="100%", fig.alt = "Add a descriptive note. Select 30 days for expiration from the dropdown. Select the scopes you need (repo). Click Generate at the bottom of this page. Keep this tab handy."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1ExTZsKDHKM0fIi0_tB80qd8cHvqjEDyNJp-OsnvZ19o/edit?slide=id.g360ed975019_0_5#slide=id.g360ed975019_0_5")
```

### Step 4: Generate and Copy Token

<p class=buttontext><input type="checkbox">  Scroll to bottom and click the green <div class="github_button">Generate token</div> button.<br/></p>
<input type="checkbox"> CRITICAL: Copy the token immediately - GitHub will only show it once (perhaps keep the window open as you copy it)<br/>
<input type="checkbox"> The token will look like: `ghp_xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx`<br/>

::: {.notice}
**Warning:** If you navigate away without copying the token, you'll need to create a new one! To create a new one, follow Steps 1-4 (within Phase 1) again.
:::

## Phase 2: Storing the Secret in Your Repository

### Step 5: Navigate to Repository Settings

<input type="checkbox"> Go to your capstone sandbox repository (not your personal settings)<br/>
<input type="checkbox"> Click the "Settings" tab at the top of the repository page<br/>
<input type="checkbox"> In the left sidebar, find "Secrets and variables"<br/>
<input type="checkbox"> Click "Actions" under "Secrets and variables"<br/>

```
Repository navigation: Repository → Settings → Secrets and variables → Actions
```

### Step 6: Create New Repository Secret

<input type="checkbox"> Click the green <div class="github_button"> New repository secret</div> button<br/>
<input type="checkbox"> Enter the secret details:<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; <input type="checkbox"> Name: `ASSIGNMENT_SECRET`<br/>
   &nbsp;&nbsp;&nbsp;&nbsp; <input type="checkbox"> Secret: Paste your personal access token from Step 4<br/>
<p class=buttontext><input type="checkbox"> Click the green <div class="github_button"> Add secret</div> button <br/></p>

::: {.warning}
Your repository secret must be named `ASSIGNMENT_SECRET` **exactly**. It is case sensitive.
:::

```{r echo=FALSE, out.width="100%", fig.alt = "Save your copied personal access token as a repository secret naming it exactly ASSIGNMENT underscore SECRET where everything is capitalized and the two words are separated by an underscore with no spaces"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1ExTZsKDHKM0fIi0_tB80qd8cHvqjEDyNJp-OsnvZ19o/edit?slide=id.g360ed975019_0_17#slide=id.g360ed975019_0_17")
```

### Step 7: Verify Secret Creation

<input type="checkbox"> You should see `ASSIGNMENT_SECRET` listed in your repository secrets<br/>
<input type="checkbox"> The value will show as "Updated X seconds ago" but won't display the actual token<br/>

If you need to update the secret, click the secret name and choose "Update"

## Phase 3: Testing Your Secret

Because you've made changes to your repository settings, but not changes to the files within the repository, we can't open a Pull Request to run an automatic test/evaluation of your work. Instead, we'll have to manually trigger the GitHub Action Evaluator Test.

In order to report success or failure, the test will open an Issue in your GitHub repo (since there is no open Pull Request to comment on).

### Step 8: Trigger the Evaluation Action

<input type="checkbox"> Navigate to the "Actions" tab in your repository <br/>
<input type="checkbox"> Look for "GitHub Secret Evaluation" in the workflows list on the left <br/>
<input type="checkbox"> Click on "GitHub Secret Evaluation" <br/>
<input type="checkbox"> Click the "Run workflow" dropdown (you may need to select branch "main") <br/>
<p class=buttontext><input type="checkbox"> Click the green <div class="github_button"> Run workflow</div> button again to confirm<br/></p>


```{r, echo=FALSE, out.width="100%", fig.alt = "Go to the Actions tab (on the top banner) to find the evaluation GitHub Action for the GitHub Secret Assignment in the left panel. From there, select the Run workflow button in the top right. Verify it is running on the main branch and click the green run workflow button."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1ExTZsKDHKM0fIi0_tB80qd8cHvqjEDyNJp-OsnvZ19o/edit?slide=id.g36aecf6d4eb_0_0#slide=id.g36aecf6d4eb_0_0")
```

### Step 9: Monitor the Evaluation & Interpret the Results

The evaluation will take a few moments to check if your secret exists and is valid. Don't worry if it takes a minute - the system needs to verify everything is configured correctly

<input type="checkbox"> Watch for the workflow to start running (you'll see a yellow circle, then green checkmark or red X)<br/>
<input type="checkbox"> Navigate to the "Issues" tab in your repository<br/>
<input type="checkbox"> Look for a new issue created by the evaluation system<br/>
<input type="checkbox"> Click on the issue to see detailed results<br/>
<input type="checkbox"> Interpret the results and proceed accordingly:<br/>
  &nbsp;&nbsp;&nbsp;&nbsp; <input type="checkbox"> If the issue title indicates success, and the description of the issue contains a validation, skip to Phase 5.<br/>
  &nbsp;&nbsp;&nbsp;&nbsp; <input type="checkbox"> If the issue title indicates failure, and the description of the issue contains error messages, visit Phase 4: Troubleshooting and Refinement of these instructions.<br/>

## Phase 4: Troubleshooting and Refinement

### Step 10: Debug Failures

#### Secret Not Found:

Verify:<br/>
<input type="checkbox"> Check the secret name: Must be exactly `ASSIGNMENT_SECRET`<br/>
<input type="checkbox"> Verify secret location:<br/>
  &nbsp;&nbsp;&nbsp;&nbsp; <input type="checkbox"> Must be in repository secrets as described in Step 6, not personal settings<br/>
  &nbsp;&nbsp;&nbsp;&nbsp; <input type="checkbox"> Secret is stored in the correct sandbox repository (not `fhdsl/capstone-sandbox` but instead your own copy)<br/>
<input type="checkbox"> Re-create if needed: Delete and recreate the secret with correct name<br/>

#### "Invalid token" Error

Verify:<br/>
<input type="checkbox"> Token was copied correctly (no extra spaces)<br/>
<input type="checkbox"> Token hasn't expired<br/>
<input type="checkbox"> Token has 'repo' scope permissions<br/>
<input type="checkbox"> Token was generated for the correct GitHub account<br/>

What to do to fix:<br/>
<input type="checkbox"> Generate a new PAT: Your token may have expired or been created incorrectly<br/>
<input type="checkbox"> Check scopes: Ensure "repo" permission is selected<br/>
<input type="checkbox"> Update the secret: Replace the old token with the new one<br/>

#### Workflow Doesn't Run

In this case, if the workflow doesn't run, it won't create an issue reporting on success/failure or you may not be able to find it to manually trigger it in Step 8.

Ensure:<br/>
<input type="checkbox"> You're triggering from the correct repository<br/>
<input type="checkbox"> You're selecting the main branch<br/>
<input type="checkbox"> The workflow file (`secret_assignment_eval.yml`) exists in `.github/workflows/`<br/>
<input type="checkbox"> You have permissions to run workflows<br/>

### Step 11: Re-running Evaluation

If you need to fix issues and re-test:<br/>
<input type="checkbox"> Make necessary corrections (recreate secret, new token, etc.)<br/>
<input type="checkbox"> Return to Actions tab<br/>
<input type="checkbox"> Run the "GitHub Secret Evaluation" workflow again<br/>
<input type="checkbox"> Check for a new issue with updated results<br/>

```{r echo=FALSE, out.width="100%", fig.alt = "To troubleshoot this assignment, you may need to regenerate a personal access token in your profile settings and recreate a token within the repository settings. After that you will visit the issues tab on that repository to check if the evaluation action ran successfully."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1ExTZsKDHKM0fIi0_tB80qd8cHvqjEDyNJp-OsnvZ19o/edit?slide=id.g36aecf6d4eb_0_5#slide=id.g36aecf6d4eb_0_5")
```

## Phase 5: Completion

::: {.certificate}

If you are taking this course on Coursera: 

**Collect Your Validation Code**

<input type="checkbox"> Once your evaluation action succeeds, copy the validation code from the Issue description. You'll need the code in order to submit your quiz.<br/>

**Submit Your Quiz on Coursera**

<input type="checkbox"> Go to your Coursera Graded Assignment ("Making a GitHub Secret")<br/>
<input type="checkbox"> Paste the validation code exactly as provided in the issue<br/>
<input type="checkbox"> Submit the quiz to get credit for your work<br/>

`r config::get("icons8")`
:::

### Clean Up (Optional but Recommended)

After completing the assignment:<br/>
<input type="checkbox"> Consider revoking the PAT if you don't plan to use it further<br/>
<input type="checkbox"> Delete the secret if it's no longer needed<br/>
<input type="checkbox"> Reflect on/Document what you learned about secret management<br/>


This assignment teaches a crucial skill for modern computational research: secure credential management. In real research environments, you'll use these same techniques to store API keys for cancer databases, authentication tokens for computing clusters, and credentials for automated research workflows. The ability to securely manage secrets enables sophisticated automation while maintaining security standards required for sensitive cancer research data.
